---
title: "What Factors Make a Masterpiece?"
author: "Noah Love"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spotifyr)
library(tidyverse)
library(knitr)
library(lubridate)
library(tidyverse)
library(corrplot)
library(ggridges)

```

## Introduction
Throughout our semester in Masterpieces of Western Music, we studied a variety of pieces and composers, and yet haven't arrived at a definitive set of characteristics that compose an individual masterpiece. Further, we compiled our own personal favorite songs, a playlist fit for if Manhattan was set adrift into the ocean and Columbia students alone chose the music we must survive on. 

### Spotify Data Analysis 
Spotify music provides an incredible API that has a wealth of information related to each and every song in their repository. Spotify themselves do the hard work of data processing individual characteristics like time signature and qualitative descriptors like loudness, instrumentalness or acousticness. There are broadly speaking 3 categories of factors:

- Mood: Danceability, Valence, Energy, Tempo
- Properties: Loudness, Speechiness, Instrumentalness
- Context: Liveness, Acousticness





```{r message=FALSE, warning=FALSE, include=FALSE}

my_id <- ''

Sys.setenv(SPOTIFY_CLIENT_ID = '')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '')

access_token <- get_spotify_access_token()
```

In our scenario, the easiest way to pull the characteristics is to use the individual playlist identifiers. 

Columbia's desert playlist: 63HV8Y11tj4Yq8vj8TYpLm

Our masterpieces: 46l09Vietu0U65GEyyWH1g

They can be easily called using:

```{r}
desert_island <- get_playlist(
    playlist_id = "63HV8Y11tj4Yq8vj8TYpLm",
    authorization = access_token
)
```

```{r include=FALSE}
tracks <- desert_island$tracks
```

```{r find track ids, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
larger_frame <- desert_island[13]

test <- larger_frame[1]$tracks

df <- test$items

tracks <- df$track.id

tracksMatrix <- as.matrix(tracks)
```

```{r get song information, message=FALSE, warning=FALSE, include=FALSE}
more_info_list <- larger_frame$tracks

more_info <- as_tibble(more_info_list$items)

larger_frame_important <- more_info %>% 
    select(track.id, track.artists, track.duration_ms, track.explicit, track.name, track.album.name, track.album.release_date, track.popularity)

```



```{r message=FALSE, warning=FALSE, include=FALSE}

track_features <- get_track_audio_features("7t86fVeDAd63ThaR0ZkxXS")
#track_features2 <- get_track_audio_analysis("79MMMdYL00iwKVHBSAUkLY")

i <- 0
datalist  <- list()

for(i in 1:length(tracks)){
    track_individuals <- get_track_audio_features(tracks[i])
    
    track_individuals$i #which iteration
    
    datalist[[i]] <- track_individuals
}

spotify_data_half <- do.call(rbind, datalist)
```

```{r include=FALSE}

# rename the song characteristics column to match larger set
spotify_data_half <- spotify_data_half %>% 
    relocate(id) %>% 
    dplyr::rename(track.id = id)
```

```{r include=FALSE}
#merge the two smaller sets
spotify_data <- full_join(spotify_data_half, larger_frame_important, by = "track.id")

```



```{r make artists tibble desert, message=FALSE, warning=FALSE, include=FALSE}
# initialize artists column
spotify_data <- spotify_data %>% 
     add_column(artists = "NAME")

i <- 0
#convert artist embedded data table into only first artist char
for(i in 1:length(spotify_data$track.artists)){
    
    # create tibble of artist and song id
    single_artist <- as.data.frame(spotify_data$track.artists[i])
    single_artist <- as.tibble(single_artist)
    
    single_artist_name <- single_artist$name

    spotify_data$artists[i] <- single_artist_name
}

```

```{r add artists in and clean data frame, include=FALSE}
spotify_data <- spotify_data %>% 
    select(-type, -uri, -track_href, -analysis_url, -track.artists) %>% 
    relocate(track.name)
    
head(spotify_data)

glimpse(spotify_data)
```

First, some simple exploration of our choices to fill our desert island with sound. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular
top_artists <- spotify_data %>%
    group_by(artists)  %>%
    dplyr::summarise(n_apperance = n()) %>%
    filter(n_apperance > 3) %>%
    arrange(desc(n_apperance))

top_artists$artists <- factor(top_artists$artists, levels = top_artists$artists[order(top_artists$n_apperance)]) # in order to visualise the list in descending order 

ggplot(top_artists, aes(x = artists, y = n_apperance)) +
    geom_bar(stat = "identity",  fill = "tomato2", width = 0.6 ) + 
    labs(title = "Top artists on our desert island", x = "Artists", y = "Number of apperances on the desert playlist") +
    theme(plot.title = element_text(size=15,hjust=-.3,face = "bold"), axis.title = element_text(size=12)) +
    geom_text(aes(label=n_apperance), hjust = 2, size = 3, color = 'white') +
    coord_flip()
```
Although some of this is related to people submitting multiple, of the 132 songs, there is a wide variety of artists in the most frequent songs. Classics, jazz, and pop all populate our most selected pieces. The recording times also vary widly, ranging from 1961 to june of this year, although as a whole the median is decently close to present. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
spotify_data$track.album.release_date <- as.Date(spotify_data$track.album.release_date)

summary(spotify_data$track.album.release_date)
```

Through the use of spotify's descriptors, we are able to look at how they interact together. Using a correlation plot on our desert island dataset, it appears very few things are closely related. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

spotify_data_num <- spotify_data[,-c((1:2),(17:19),(21))] 

mtCor <- cor(spotify_data_num)
corrplot(mtCor, method = "ellipse", type = "upper", tl.srt = 45)

```

Energy and loudness is strongly correlated. Acousticness and instrumentalness are strongly negatively correlated with energy and loudness. This seems reasonable as there is very infrequently really loud instrumental or acappella songs. 

```{r}
spotify_data %>% 
  arrange(-acousticness) %>% 
  select(track.name, acousticness) %>% 
  head(4) %>% 
  kable() 
  
```
Particularly pulled out in this "acousticness" are piano pieces. At the top is Debussy in a rather slow and tender Prelude. 

Interestingly, popularity (to all spotify listeners) is not very correlated with anything except loudness, and negatively to acousticness. It appears we like large, grand sounds. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular

spotify_data$key <- as.character(spotify_data$key)
spotify_data$key <- plyr::revalue(spotify_data$key, c("0" = "C", "1" = "C♯,D♭", "2" = "D", "3" = "D♯,E♭", "4" = "E", "5" =  "F", "6" = "F♯,G♭","7" = "G","8" = "G♯,A♭","9" = "A","10" = "A♯,B♭","11" = "B"))

song_keys <- spotify_data %>%
    dplyr::group_by(key) %>%
    dplyr::summarise(n_key = n()) %>%
    arrange(desc(n_key))

song_keys$key <- factor(song_keys$key, levels = song_keys$key[order(song_keys$n_key)]) # in order to visualise the keys in descending order

ggplot(song_keys, aes(x = reorder(key,-n_key), y = n_key, fill = reorder(key,-n_key))) +
    geom_bar(stat = "identity") +
    labs(title = "Distribution of the Keys of Top Songs", x = "Keys", y = "Count of Keys on the Top 100") +
    geom_text(aes(label=n_key), position = position_stack(vjust = 0.8)) +
    theme_bw() +
    theme(plot.title = element_text(size=15,face = "bold"), axis.title = element_text(size=12)) +
    theme(legend.position="none")
```



```{r echo=FALSE, message=FALSE, warning=FALSE}

# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular
correlated_density <- ggplot(spotify_data) +
    geom_density(aes(energy, fill ="energy", alpha = 0.1)) + 
    geom_density(aes(valence, fill ="valence", alpha = 0.1)) + 
    geom_density(aes(danceability, fill ="danceability", alpha = 0.1)) + 
    scale_x_continuous(name = "Energy, Valence and Danceability") +
    scale_y_continuous(name = "Density") +
    ggtitle("Density plot of Energy, Valence and Danceability") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          text = element_text(size = 12)) +
    theme(legend.title=element_blank()) +
    scale_fill_brewer(palette="Accent")

correlated_density
```



```{r echo=FALSE, message=FALSE, warning=FALSE}

# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular


loudness_density <- ggplot(spotify_data) +
    geom_density(aes(loudness, fill ="loudness")) + 
    scale_x_continuous(name = "Loudness") +
    scale_y_continuous(name = "Density") +
    ggtitle("Density plot of Loudness") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
            text = element_text(size = 12)) +
    theme(legend.title=element_blank()) +
    scale_fill_brewer(palette="Paired")

print(loudness_density)
```


Adagio – slowly with great expression[10] (66–76 bpm)
Andante – at a walking pace (76–108 bpm)
Moderato – at a moderate speed (108–120 bpm)
Allegro – fast, quickly, and bright (120–156 bpm)
Vivace – lively and fast (156–176 bpm)
Presto – very, very fast (168–200 bpm)



```{r message=FALSE, warning=FALSE, include=FALSE}
joy <- get_artist_audio_features('Hector Berlioz')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Symphonie fantastique, Op. 14, H. 48: I. Rêveries - Passions	
# Symphonie fantastique, Op. 14, H. 48: II. Un bal	
# Symphonie fantastique, Op. 14, H. 48: III. Scène aux champs	
# Symphonie fantastique, Op. 14, H. 48: IV. Marche au supplice	
# Symphonie fantastique, Op. 14, H. 48: V. Songe d'une nuit de sabbat

berlioz <- joy %>% 
  filter(album_release_date == "2021-07-23")

berlioz

berlioz %>% 
    arrange(-valence) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-loudness) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-liveness) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-liveness) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-danceability) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

```




## What makes a master 

```{r echo=FALSE, message=FALSE, warning=FALSE}
masterpieces <- get_playlist(
    playlist_id = "46l09Vietu0U65GEyyWH1g",
    authorization = access_token
)

tracks_masterpieces <- masterpieces$tracks

larger_frame_masterpieces <- masterpieces[13]
test_masterpieces <- larger_frame_masterpieces[1]$tracks
df_masterpieces <- test_masterpieces$items

tracks_masterpieces <- df_masterpieces$track.id

tracksMatrix_masterpieces <- as.matrix(tracks_masterpieces)

more_info_list_masterpieces <- larger_frame_masterpieces$tracks

more_info_masterpieces <- as_tibble(more_info_list_masterpieces$items)

larger_frame_important_masterpieces <- more_info_masterpieces %>% 
    select(track.id, track.artists, track.duration_ms, track.explicit, track.name, track.album.name, track.album.release_date, track.popularity)

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
track_features_masterpieces <- get_track_audio_features("4yg4OeZxDSqyXx6NUxtxHR")

i <- 0
datalist_masterpieces  <- list()

for(i in 1:length(tracks_masterpieces)){
    track_individuals_masterpieces <- get_track_audio_features(tracks_masterpieces[i])
    
    track_individuals_masterpieces$i #which iteration
    
    datalist_masterpieces[[i]] <- track_individuals_masterpieces
}

spotify_data_half_masterpieces <- do.call(rbind, datalist_masterpieces)


# rename the song characteristics column to match larger set
spotify_data_half_masterpieces <- spotify_data_half_masterpieces %>% 
    relocate(id) %>% 
    dplyr::rename(track.id = id)

#merge the two smaller sets
spotify_data_masterpieces <- full_join(spotify_data_half_masterpieces, larger_frame_important_masterpieces, by = "track.id")

```


```{r make artists tibble masterpieces, echo=FALSE, message=FALSE, warning=FALSE}
# initialize artists column
spotify_data_masterpieces <- spotify_data_masterpieces %>% 
     add_column(artists = "NAME")

i <- 0
#convert artist embedded data table into only first artist char
for(i in 1:length(spotify_data_masterpieces$track.artists)){
    
    # create tibble of artist and song id
    single_artist_masterpieces <- as.data.frame(spotify_data_masterpieces$track.artists[i])
    single_artist_masterpieces <- as.tibble(single_artist_masterpieces)
    
    single_artist_name_masterpieces <- single_artist_masterpieces$name

    spotify_data_masterpieces$artists[i] <- single_artist_name_masterpieces
}

spotify_data_masterpieces <- spotify_data_masterpieces %>% 
    select(-type, -uri, -track_href, -analysis_url, -track.artists) %>% 
    relocate(track.name)

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
multiple_masters <- spotify_data_masterpieces %>% 
  group_by(artists) %>% 
  filter(n()>1)

ggplot(
    multiple_masters, 
    aes(x = valence, y = artists)
    ) + 
  geom_density_ridges_gradient(show.legend = FALSE) +
geom_density_ridges() + 
theme_ridges() +
labs(title = "Valence across the 'masters'")

```




