---
title: "Analysis-music-hum"
author: "Noah Love"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spotifyr)
library(tidyverse)
library(knitr)
library(lubridate)
library(tidyverse)
library(corrplot)
library(ggridges)

```

## R Markdown

Hello this is a test

```{r include=FALSE}

my_id <- ''

Sys.setenv(SPOTIFY_CLIENT_ID = '')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '')

access_token <- get_spotify_access_token()
```


```{r}
desert_island <- get_playlist(
    playlist_id = "63HV8Y11tj4Yq8vj8TYpLm",
    authorization = access_token
)
```

```{r}
tracks <- desert_island$tracks
```

```{r find track ids, echo=FALSE, message=FALSE, warning=FALSE}
larger_frame <- desert_island[13]

test <- larger_frame[1]$tracks

df <- test$items

tracks <- df$track.id

tracksMatrix <- as.matrix(tracks)
```

```{r get song information, message=FALSE, warning=FALSE}
more_info_list <- larger_frame$tracks

more_info <- as_tibble(more_info_list$items)

larger_frame_important <- more_info %>% 
    select(track.id, track.artists, track.duration_ms, track.explicit, track.name, track.album.name, track.album.release_date, track.popularity)

```



```{r message=FALSE, warning=FALSE}

track_features <- get_track_audio_features("7t86fVeDAd63ThaR0ZkxXS")
#track_features2 <- get_track_audio_analysis("79MMMdYL00iwKVHBSAUkLY")

i <- 0
datalist  <- list()

for(i in 1:length(tracks)){
    track_individuals <- get_track_audio_features(tracks[i])
    
    track_individuals$i #which iteration
    
    datalist[[i]] <- track_individuals
}

spotify_data_half <- do.call(rbind, datalist)
```

```{r}
spotify_data_half
# rename the song characteristics column to match larger set
spotify_data_half <- spotify_data_half %>% 
    relocate(id) %>% 
    dplyr::rename(track.id = id)
```

```{r}
#merge the two smaller sets
spotify_data <- full_join(spotify_data_half, larger_frame_important, by = "track.id")

```



```{r make artists tibble, message=FALSE, warning=FALSE}
# initialize artists column
spotify_data <- spotify_data %>% 
     add_column(artists = "NAME")

i <- 0
#convert artist embedded data table into only first artist char
for(i in 1:length(spotify_data$track.artists)){
    
    # create tibble of artist and song id
    single_artist <- as.data.frame(spotify_data$track.artists[i])
    single_artist <- as.tibble(single_artist)
    
    single_artist_name <- single_artist$name

    spotify_data$artists[i] <- single_artist_name
}

```

```{r add artists in and clean data frame}
spotify_data <- spotify_data %>% 
    select(-type, -uri, -track_href, -analysis_url, -track.artists) %>% 
    relocate(track.name)
    
```



```{r}
# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular
top_artists <- spotify_data %>%
    group_by(artists)  %>%
    dplyr::summarise(n_apperance = n()) %>%
    filter(n_apperance > 3) %>%
    arrange(desc(n_apperance))

top_artists$artists <- factor(top_artists$artists, levels = top_artists$artists[order(top_artists$n_apperance)]) # in order to visualise the list in descending order 

ggplot(top_artists, aes(x = artists, y = n_apperance)) +
    geom_bar(stat = "identity",  fill = "tomato2", width = 0.6 ) + 
    labs(title = "Top Artists on our desert island", x = "Artists", y = "Number of apperances on the desert playlist") +
    theme(plot.title = element_text(size=15,hjust=-.3,face = "bold"), axis.title = element_text(size=12)) +
    geom_text(aes(label=n_apperance), hjust = 2, size = 3, color = 'white') +
    coord_flip()
```


```{r}


spotify_data

spotify_data_num <- spotify_data[,-c((1:2),(17:19),(21))] 
 
spotify_data_num

mtCor <- cor(spotify_data_num)
corrplot(mtCor, method = "ellipse", type = "upper", tl.srt = 45)

```

Energy and loudness is strongly correlated. Acousticness and instrumentalness are strongly negatively correlated with energy and loudness. 

Interestingly, popularity (to all spotify listeners) is not very correlated with anything except loudness, and negatively to acousticness. It appears we like large, grand sounds. 

```{r}

# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular

spotify_data$key <- as.character(spotify_data$key)
spotify_data$key <- plyr::revalue(spotify_data$key, c("0" = "C", "1" = "C♯,D♭", "2" = "D", "3" = "D♯,E♭", "4" = "E", "5" =  "F", "6" = "F♯,G♭","7" = "G","8" = "G♯,A♭","9" = "A","10" = "A♯,B♭","11" = "B"))

song_keys <- spotify_data %>%
    dplyr::group_by(key) %>%
    dplyr::summarise(n_key = n()) %>%
    arrange(desc(n_key))

song_keys$key <- factor(song_keys$key, levels = song_keys$key[order(song_keys$n_key)]) # in order to visualise the keys in descending order

ggplot(song_keys, aes(x = reorder(key,-n_key), y = n_key, fill = reorder(key,-n_key))) +
    geom_bar(stat = "identity") +
    labs(title = "Distribution of the Keys of Top Songs", x = "Keys", y = "Count of Keys on the Top 100") +
    geom_text(aes(label=n_key), position = position_stack(vjust = 0.8)) +
    theme_bw() +
    theme(plot.title = element_text(size=15,face = "bold"), axis.title = element_text(size=12)) +
    theme(legend.position="none")
```



```{r}

# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular
correlated_density <- ggplot(spotify_data) +
    geom_density(aes(energy, fill ="energy", alpha = 0.1)) + 
    geom_density(aes(valence, fill ="valence", alpha = 0.1)) + 
    geom_density(aes(danceability, fill ="danceability", alpha = 0.1)) + 
    scale_x_continuous(name = "Energy, Valence and Danceability") +
    scale_y_continuous(name = "Density") +
    ggtitle("Density plot of Energy, Valence and Danceability") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          text = element_text(size = 12)) +
    theme(legend.title=element_blank()) +
    scale_fill_brewer(palette="Accent")

correlated_density
```



```{r}

# https://www.kaggle.com/cihanoklap/top-songs-on-spotify-what-makes-them-popular


loudness_density <- ggplot(spotify_data) +
    geom_density(aes(loudness, fill ="loudness")) + 
    scale_x_continuous(name = "Loudness") +
    scale_y_continuous(name = "Density") +
    ggtitle("Density plot of Loudness") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
            text = element_text(size = 12)) +
    theme(legend.title=element_blank()) +
    scale_fill_brewer(palette="Paired")

print(loudness_density)
```


Adagio – slowly with great expression[10] (66–76 bpm)
Andante – at a walking pace (76–108 bpm)
Moderato – at a moderate speed (108–120 bpm)
Allegro – fast, quickly, and bright (120–156 bpm)
Vivace – lively and fast (156–176 bpm)
Presto – very, very fast (168–200 bpm)



```{r}
joy <- get_artist_audio_features('Hector Berlioz')
joy
```

```{r}
# Symphonie fantastique, Op. 14, H. 48: I. Rêveries - Passions	
# Symphonie fantastique, Op. 14, H. 48: II. Un bal	
# Symphonie fantastique, Op. 14, H. 48: III. Scène aux champs	
# Symphonie fantastique, Op. 14, H. 48: IV. Marche au supplice	
# Symphonie fantastique, Op. 14, H. 48: V. Songe d'une nuit de sabbat

berlioz <- joy %>% 
  filter(album_release_date == "2021-07-23")

berlioz

berlioz %>% 
    arrange(-valence) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-loudness) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-liveness) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-liveness) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

berlioz %>% 
    arrange(-danceability) %>% 
    select(.data$track_name, .data$valence) %>% 
    head(5) %>% 
    kable()

```





```{r}
ggplot(berlioz, aes(x = energy,y=fct_rev(track_name))) + 
  geom_density_ridges(scale = 2, size = 0.1, rel_min_height = 0.03,quantile_lines=TRUE,quantiles=2) +
  theme_ridges()+
  scale_x_continuous( expand = c(0, 0)) +
  labs(title= "Foo Fighters energy by album",subtitle = "Based on data from Spotify", y="Album", x = "Energy")
```

```{r}
ggplot(
    berlioz, 
    aes(x = liveness, y = track_name)
    ) + 
geom_density_ridges() + 
theme_ridges() +
labs(title = "Joyplot of Joy Division's joy distributions")
```

## What makes a master 

```{r}
masterpieces <- get_playlist(
    playlist_id = "46l09Vietu0U65GEyyWH1g",
    authorization = access_token
)

tracks_masterpieces <- masterpieces$tracks

larger_frame_masterpieces <- masterpieces[13]
test_masterpieces <- larger_frame_masterpieces[1]$tracks
df_masterpieces <- test_masterpieces$items

tracks_masterpieces <- df_masterpieces$track.id

tracksMatrix_masterpieces <- as.matrix(tracks_masterpieces)

more_info_list_masterpieces <- larger_frame_masterpieces$tracks

more_info_masterpieces <- as_tibble(more_info_list_masterpieces$items)

larger_frame_important_masterpieces <- more_info_masterpieces %>% 
    select(track.id, track.artists, track.duration_ms, track.explicit, track.name, track.album.name, track.album.release_date, track.popularity)

larger_frame_important_masterpieces
```



```{r message=FALSE, warning=FALSE}
track_features_masterpieces <- get_track_audio_features("4yg4OeZxDSqyXx6NUxtxHR")

i <- 0
datalist_masterpieces  <- list()

for(i in 1:length(tracks_masterpieces)){
    track_individuals_masterpieces <- get_track_audio_features(tracks_masterpieces[i])
    
    track_individuals_masterpieces$i #which iteration
    
    datalist_masterpieces[[i]] <- track_individuals_masterpieces
}

spotify_data_half_masterpieces <- do.call(rbind, datalist_masterpieces)


# rename the song characteristics column to match larger set
spotify_data_half_masterpieces <- spotify_data_half_masterpieces %>% 
    relocate(id) %>% 
    dplyr::rename(track.id = id)

#merge the two smaller sets
spotify_data_masterpieces <- full_join(spotify_data_half_masterpieces, larger_frame_important_masterpieces, by = "track.id")

```


```{r make artists tibble, message=FALSE, warning=FALSE}
# initialize artists column
spotify_data_masterpieces <- spotify_data_masterpieces %>% 
     add_column(artists = "NAME")

i <- 0
#convert artist embedded data table into only first artist char
for(i in 1:length(spotify_data_masterpieces$track.artists)){
    
    # create tibble of artist and song id
    single_artist_masterpieces <- as.data.frame(spotify_data_masterpieces$track.artists[i])
    single_artist_masterpieces <- as.tibble(single_artist_masterpieces)
    
    single_artist_name_masterpieces <- single_artist_masterpieces$name

    spotify_data_masterpieces$artists[i] <- single_artist_name_masterpieces
}

spotify_data_masterpieces <- spotify_data_masterpieces %>% 
    select(-type, -uri, -track_href, -analysis_url, -track.artists) %>% 
    relocate(track.name)

spotify_data_masterpieces
    
```



```{r}
multiple_masters <- spotify_data_masterpieces %>% 
  group_by(artists) %>% 
  filter(n()>1)

ggplot(
    multiple_masters, 
    aes(x = valence, y = artists)
    ) + 
  geom_density_ridges_gradient(show.legend = FALSE) +
geom_density_ridges() + 
theme_ridges() +
labs(title = "Valence across the 'masters'")

```



